Asteroseismology of GD358 with complex C/O core profiles

### 2016-10-01 msnew.tex
added Andy
some changes from ref report

\author {Agn\`es~Bischoff-Kim\altaffilmark{1},
J.~L. Provencal\altaffilmark{2,3},
M.~H.~Montgomery\altaffilmark{3,4},
H.~L. Shipman\altaffilmark{2,3},
Samuel~T.~Harrold\altaffilmark{4},
B.~Howard\altaffilmark{3},
W.~Strickland\altaffilmark{5},
D.~Chandler\altaffilmark{5},
D.~Campbell\altaffilmark{5},
A.~Arredondo\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~P.~Russell\altaffilmark{5},
D.~Doyle\altaffilmark{5},
A.~Brickhouse\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~Peters\altaffilmark{5},
P.~A. Bradley\altaffilmark{6},
S.-L. Kim\altaffilmark{7},
X.~J.~Jiang\altaffilmark{8},
Y-N.~Mao\altaffilmark{8},
A.~V.~Kusakin\altaffilmark{9},
A.~V.~Sergeev\altaffilmark{10,11},
M.~Andreev\altaffilmark{10,11},
A.~Maksim\altaffilmark{10,11},
S. Velichko\altaffilmark{12},
R.~Janulis\altaffilmark{12},
E.~Pakstiene\altaffilmark{12},
F.~Ali\c cavu\c s\altaffilmark{13},
N. Horoz\altaffilmark{13},
S.~Zola\altaffilmark{14,15},
W.~Og{\l}oza\altaffilmark{14,15},
D.~Koziel-Wierzbowska\altaffilmark{14,15},
T.~Kundera\altaffilmark{14,15},
D. Jableka\altaffilmark{14,15},
B. Debski\altaffilmark{14,15},
S. Meingast\altaffilmark{16},
T. Nagel\altaffilmark{17},
L. Loebling\altaffilmark{17},
C. Heinitz\altaffilmark{17},
D. Hoyer\altaffilmark{17},
Zs. Bogn\'ar\altaffilmark{18}
}
\altaffiltext{1}{Penn State Worthington Scranton, Dunmore, PA 18512; axk55@psu.edu}
\altaffiltext{2}{University of Delaware, Department of Physics and Astronomy
Newark, DE 19716; jlp@udel.edu}
\altaffiltext{3}{Delaware Asteroseismic Research Center, Mt. Cuba Observatory,
Greenville, DE 19807}
\altaffiltext{4}{Department of Astronomy, University of Texas, Austin, TX 78712;
mikemon@rocky.as.utexas.edu}
\altaffiltext{5}{Meyer Observatory and Central Texas Astronomical Society, 209
Paintbrush, Waco, TX 76705; chandler@vvm.com}
\altaffiltext{7}{Korea Astronomy and Space Science Institute, Daejeon 34055, Korea}
\altaffiltext{8}{National Astronomical Observatories, Academy of Sciences, Beijing 10012, People's Republic of
China; xjjiang@bao.ac.cn}
\altaffiltext{9}{Fesenkov Astrophysical Institute, Almaty 050020, Kazakhstan}
\altaffiltext{10}{Ukrainian National Academy of Sciences, International Center for Astronomical,
Medical and Ecological Research, 27, Akademika Zabolotnoho Ave., 03680 Kyiv, Ukraine; sergeev@terskol.com }
\altaffiltext{11}{Russian Academy of Sciences, Institute of Astronomy, Terskol Branch, 81, Elbrus Ave., ap. 33, Tyrnyauz,
Kabardino-Balkaria Republic, 361623, Russian Federation; sergeev@terskol.com}
\altaffiltext{12}{Institute of Theoretical Physics and Astronomy, Vilnius
University, Vilnius, Lithuania; jr@itpa.lt}
\altaffiltext{13}{Ulupinar Observatory, \c Canakkale Onsekiz Mart University, Turkey}
\altaffiltext{14}{Astronomical Observatory of the Jagiellonian University, ul. Orla
171, 30-244 Cracow, Poland; szola@oa.uj.edu.pl}
\altaffiltext{15}{Mount Suhora Observatory, Cracow Pedagogical University, Ul.
Podchorazych 2, 30-084 Krakow, Poland; zola@astro1.as.ap.krakow.pl}
\altaffiltext{16}{Institut f\H ur Astrophysik, Universit\H at Wien, T\H urkenschanzstrasse 17, 1180 Wien, Austria}
\altaffiltext{17}{Institut fuer Astronomie und Astrophysik, Kepler Center for Astro and Particle Physics,
Eberhard Karls Universitaet Tuebingen, Sand 1, 72076 Tuebingen Germany; nagel@astro.uni-tuebingen.de}
\altaffiltext{18}{Konkoly Observatory, MTA CSFK, Konkoly Thege M. u. 15-17, H-1121 Budapest, Hungary}

@@

Data reduction follows the steps described in \citet{Provencal12}. In brief, the new observations 
were obtained with different CCD photometers, each with distinct effective bandpasses. 
We minimize the bandpass issues by using CCDs with similar detectors whenever possible and 
employing a red cutoff filter (BG40 or S8612) to normalize wavelength response and 
reduce extinction effects. We corrected each image for bias and thermal noise, and 
normalized by the flat field. Aperture photometry using the Maestro photometry pipeline 
described by \citet{Dalessio10} was performed on each image, utilizing a range of 
aperture sizes for the target and comparison stars. For each individual nightly run, we 
chose the combination of aperture size and comparison star(s) producing the highest 
signal/noise light curve.

We used the WQED pipeline \citep{wqed} to examine each nightly light curve for photometric 
quality, remove outlying points, divide by suitable comparison stars, and correct for 
differential extinction. Our observational technique is therefore not sensitive to 
oscillations longer than a few hours. The result is a series of light curves 
with amplitude variations represented as fractional intensity (mmi). The unit is a linear 
representation of the fractional intensity of modulation (1 mmi $\approx$ 1 mmag). We present
our Fourier transforms (FTs) in units of modulation amplitude 
($1\ {\rm mma}=1 \times 10^{-3}\ {\rm ma} = 0.1 \% = 1\ {\rm ppt}$).

The final reduction step is to combine the individual light curves 
(an example is shown in Fig. \ref{tersk}) to produce complete light curves for GD358 for each 
observing season. We assume GD358 oscillates about a mean light level. This assumption 
allows us to carefully assess overlapping segments from different telescopes and identify 
and correct any vertical offsets. As discussed in detail in \citet{Provencal09}, we find 
no significant differences between the noise levels of amplitude spectra using: 
1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

@@

The complexities associated with GD358's pulsations (see Section \ref{freq}) led us to also 
re-reduce all available archival data \citep{Provencal09, Bradley04, Kepler03, Winget94, Winget82} to
insure continuity in methodology. The final result is a series of light curves 
for each observing season between 1982 and 2016. 
For the new observations, 2007 contains 8.1 hrs of observation, 2008 26.1 hrs, 2009 45 hrs, 
2010 201.3 hrs, 2011 401.1 hrs, 2012 150.5 hrs, 2013 87 hrs, 2014 184.6 hrs, 2015 55 hrs and 2016 90.6 hrs. 
Our coverage is not complete, and this incompleteness produces spectral leakage in the 
amplitude spectra. To quantify this, we sampled single sinusoids using exactly the 
same times as the original data for each season. The resulting amplitude spectrum, 
or ``spectral window'', is the pattern produced in the FT by a single frequency. 
The FTs for 2010, 2011, and 2014 are given in Fig.~\ref{fts}.

@@

We use {\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
select statistically significant peaks in each FT. We adopt the criterion that a peak must have 
an amplitude at least four times greater than the average noise level in the given frequency range 
\citep{Provencal12}. We define ``noise'' as the frequency-dependent 
average amplitude after removal of the dominant frequencies. This is a conservative estimate, 
as it is impossible to ensure that all of the ``real'' frequencies are removed when determining
the noise level. This is certainly true for GD358, where the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies (see Section \ref{mode}, \citet{Provencal09}).  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2016 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. To confirm our uncertainty 
estimates, we calculated Monte Carlo simulations using the routine provided in {\sl Period04}.  
This routine generates a set of light curves using the original times, the fitted frequencies and 
amplitudes, and added Gaussian noise. A least squares fit is performed on each simulated light 
curve, with the distribution of fit parameters giving the uncertainties. Our Monte Carlo results 
are consistent with our average noise estimates.

@@

Having established our baseline noise levels, we proceed to frequency selection and identification.
Our frequency selection procedure involves identifying the largest amplitude resolved peak in the FT,
fitting a sinusoid at that frequency to the data set, subtracting the fit from the light curve, 
recomputing the FT, examining the residuals, and repeating the process until no significant power 
remains. This technique, known as prewhitening, must be employed with an over-abundance of caution, 
especially since we are aware of amplitude and/or frequency modulation in our data set. A detailed
discussion of the prewhitening procedure and steps taken to minimize the effects of amplitude 
modulation is given in \citet{Provencal09}. Our final identifications of independent frequencies
detected in each observing season are given in Table \ref{freq1} and Table \ref{freq2}.

@@

Fig.~\ref{schematic} presents a schematic representation of GD358's 
independent frequencies detected between 1982 and 2016.  
The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret the bands to represent a series of modes of radial overtone $k$. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of stable multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands (Fig.~\ref{rot}). This does not imply that the 
structures found by \cite{Winget94}, \cite{Kepler03}, \cite{Montgomery10} and others are not real. In 
particular, it is clear that the $k=12$ mode can be explained by the oblique rotation model during 2006. However, 
previous work represents short snapshots of GD358. When examined over timescales of decades, these structures 
do not regularly recur, and so cannot be interpreted as simple rotational splitting. The lower frequency band widths 
are most likely the result of various processes producing amplitude and/or frequency modulation that obscures any 
underlying signatures of rotational splitting. 
\begin{figure}
 \epsscale{1.0}
 \plotone{gd358k17_k8.eps}
% \plottwo{gd358multifk17.eps}{gd358multifk8.eps}
 \caption{The detailed schematic distribution of frequencies for the bands near 1300({\em{right}}
 and 2360 \muHz\ ({\em{left}}). As an example of the higher frequency bands, the 1300 \muHz\ band 
 shows no multiplet structure, unlike the 2360 \muHz\ band. Here we see clear evidence of multiplet 
 structure. Assuming the triplet represents
 $\ell=1$, this implies a rotation period of $\approx 1.5$ days. Note changes in the x scale for both axes.
 (A color version of this figure is available in the online journal.)
 \label{rot}}
\end{figure}

>>>>

\author {Agn\`es~Bischoff-Kim\altaffilmark{1},
J.~L. Provencal\altaffilmark{2,3},
M.~H.~Montgomery\altaffilmark{3,4},
H.~L. Shipman\altaffilmark{2,3},
Samuel~T.~Harrold\altaffilmark{4},
B.~Howard\altaffilmark{3},
W.~Strickland\altaffilmark{5},
D.~Chandler\altaffilmark{5},
D.~Campbell\altaffilmark{5},
A.~Arredondo\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~P.~Russell\altaffilmark{5},
D.~Doyle\altaffilmark{5},
A.~Brickhouse\altaffilmark{5},
R.~Linn\altaffilmark{5},
D.~Peters\altaffilmark{5},
P.~A. Bradley\altaffilmark{6},
S.-L. Kim\altaffilmark{7},
X.~J.~Jiang\altaffilmark{8},
Y-N.~Mao\altaffilmark{8},
A.~V.~Kusakin\altaffilmark{9},
A.~V.~Sergeev\altaffilmark{10,11},
M.~Andreev\altaffilmark{10,11},
A.~Maksim\altaffilmark{10,11},
S. Velichko\altaffilmark{12},
R.~Janulis\altaffilmark{12},
E.~Pakstiene\altaffilmark{12},
F.~Ali\c cavu\c s\altaffilmark{13},
N. Horoz\altaffilmark{13},
S.~Zola\altaffilmark{14,15},
W.~Og{\l}oza\altaffilmark{14,15},
D.~Koziel-Wierzbowska\altaffilmark{14,15},
T.~Kundera\altaffilmark{14,15},
D. Jableka\altaffilmark{14,15},
B. Debski\altaffilmark{14,15},
A. Baran\altaffilmark{15},
S. Meingast\altaffilmark{16},
T. Nagel\altaffilmark{17},
L. Loebling\altaffilmark{17},
C. Heinitz\altaffilmark{17},
D. Hoyer\altaffilmark{17},
Zs. Bogn\'ar\altaffilmark{18}
}
\altaffiltext{1}{Penn State Worthington Scranton, Dunmore, PA 18512; axk55@psu.edu}
\altaffiltext{2}{University of Delaware, Department of Physics and Astronomy
Newark, DE 19716; jlp@udel.edu}
\altaffiltext{3}{Delaware Asteroseismic Research Center, Mt. Cuba Observatory,
Greenville, DE 19807}
\altaffiltext{4}{Department of Astronomy, University of Texas, Austin, TX 78712;
mikemon@rocky.as.utexas.edu}
\altaffiltext{5}{Meyer Observatory and Central Texas Astronomical Society, 209
Paintbrush, Waco, TX 76705; chandler@vvm.com}
\altaffiltext{7}{Korea Astronomy and Space Science Institute, Daejeon 34055, Korea}
\altaffiltext{8}{National Astronomical Observatories, Academy of Sciences, Beijing 10012, People's Republic of
China; xjjiang@bao.ac.cn}
\altaffiltext{9}{Fesenkov Astrophysical Institute, Almaty 050020, Kazakhstan}
\altaffiltext{10}{Ukrainian National Academy of Sciences, International Center for Astronomical,
Medical and Ecological Research, 27, Akademika Zabolotnoho Ave., 03680 Kyiv, Ukraine; sergeev@terskol.com }
\altaffiltext{11}{Russian Academy of Sciences, Institute of Astronomy, Terskol Branch, 81, Elbrus Ave., ap. 33, Tyrnyauz,
Kabardino-Balkaria Republic, 361623, Russian Federation; sergeev@terskol.com}
\altaffiltext{12}{Institute of Theoretical Physics and Astronomy, Vilnius
University, Vilnius, Lithuania; jr@itpa.lt}
\altaffiltext{13}{Ulupinar Observatory, \c Canakkale Onsekiz Mart University, Turkey}
\altaffiltext{14}{Astronomical Observatory of the Jagiellonian University, ul. Orla
171, 30-244 Cracow, Poland; szola@oa.uj.edu.pl}
\altaffiltext{15}{Mount Suhora Observatory, Cracow Pedagogical University, Ul.
Podchorazych 2, 30-084 Krakow, Poland; zola@astro1.as.ap.krakow.pl}
\altaffiltext{16}{Institut f\H ur Astrophysik, Universit\H at Wien, T\H urkenschanzstrasse 17, 1180 Wien, Austria}
\altaffiltext{17}{Institut fuer Astronomie und Astrophysik, Kepler Center for Astro and Particle Physics,
Eberhard Karls Universitaet Tuebingen, Sand 1, 72076 Tuebingen Germany; nagel@astro.uni-tuebingen.de}
\altaffiltext{18}{Konkoly Observatory, MTA CSFK, Konkoly Thege M. u. 15-17, H-1121 Budapest, Hungary}

@@

Data reduction follows the steps described in \citet{Provencal12}. In brief, the new observations 
were obtained with CCD photometers at multiple sites, each with distinct effective bandpasses. 
We reduce the bandpass issues by using CCDs with similar detectors whenever possible and 
employing a red cutoff filter (BG40 or S8612) to normalize wavelength response and 
minimize extinction effects. We corrected each image for bias and thermal noise, and 
normalized by the flat field. Aperture photometry using the Maestro photometry pipeline 
described by \citet{Dalessio10} was performed on each image, covering a range of 
aperture sizes for the target and comparison stars. For each individual nightly run, we 
chose the combination of aperture size and comparison star(s) producing the highest 
signal/noise light curve.

We used the WQED pipeline \citep{wqed} to examine each light curve for photometric 
quality, remove outlying points, divide by suitable comparison stars, and correct for 
differential extinction. This observational technique is therefore not sensitive to 
oscillations longer than a few hours. The final product is a series of light curves 
with amplitude variations represented as fractional intensity (mmi). The unit is a linear 
representation of the fractional intensity of modulation (1 mmi $\approx$ 1 mmag). We present
our Fourier transforms (FTs) in units of modulation amplitude 
($1\ {\rm mma}=1 \times 10^{-3}\ {\rm ma} = 0.1 \% = 1\ {\rm ppt}$).

The final reduction step is to combine the individual light curves 
(an example is shown in Fig. \ref{tersk}) to produce complete light curves for GD358 for each 
observing season. As we do for all white dwarf pulators, we assume GD358 oscillates around 
a mean light level. This assumption allows us to assess overlapping light curves from multiple 
telescopes and identify and correct any vertical offsets. As discussed in detail in \citet{Provencal09}, we find 
no significant differences between the noise levels of amplitude spectra using: 
1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

@@

The complexities associated with GD358's pulsations (see Section \ref{freq}) led us to also 
re-reduce all available archival data \citep{Provencal09, Bradley04, Kepler03, Winget94, Winget82} to
insure continuity in methodology. The final result is a series of light curves 
for each observing season between 1982 and 2016. 
For the new observations, 2007 contains 8.1 hrs of observation, 2008 26.1 hrs, 2009 45 hrs, 
2010 201.3 hrs, 2011 401.1 hrs, 2012 150.5 hrs, 2013 87 hrs, 2014 184.6 hrs, 2015 55 hrs and 2016 90.6 hrs. 
Our coverage is not complete, and this incompleteness produces spectral leakage in the 
amplitude spectra. To quantify this, our standard procedure samples single sinusoids using exactly the 
same times as the original data for each season. The resulting amplitude spectrum, 
or ``spectral window'', is the pattern produced in the FT by a single frequency. 
The FTs for 2010, 2011, and 2014 are given in Fig.~\ref{fts}.

@@

We use {\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
identify statistically significant peaks in each FT. We adopt the criterion that a peak must have 
an amplitude at least four times that of the average noise level in the given frequency range 
\citep{Provencal12}. This criterion places a 99.9\% probability that the peak represents a real signal,
and is not a result of random noise \citep{Scargle82, Provencal12}.  We define ``noise'' as the frequency-dependent 
average amplitude after prewhitening of the dominant frequencies. This is unquestionably a conservative estimate, 
as it is impractical to assume that all ``real'' frequencies are removed when determining
the noise level. This is certainly true for GD358, where amplitude modulation is present, and the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies (see Section \ref{mode}, \citet{Provencal09}).  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2016 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. We also calculated Monte Carlo 
simulations using the routine provided in {\sl Period04}.  This routine generates a set of light curves using 
the original times, the fitted frequencies and amplitudes, and added Gaussian noise. A least squares fit is 
performed on each simulated light curve, with the distribution of fit parameters giving the uncertainties. 
Our Monte Carlo results are consistent with our average noise estimates.

@@

Having established our baseline noise levels, we proceed to frequency selection and identification.
Our standard frequency selection procedure involves identifying the largest amplitude resolved peak in the FT,
fitting a sinusoid with that frequency to the data set, subtracting the fit from the light curve, 
recomputing the FT, examining the residuals, and repeating the process until no significant power 
remains. This technique is known as prewhitening. Prewhitening has inherent dangers, and must be 
employed with extreme vigilance, especially as we are aware of amplitude and/or frequency 
modulation in our data set. A detailed discussion of the prewhitening procedure and steps 
taken to minimize the effects of amplitude modulation is given in \citet{Provencal09}. Our final 
identifications of independent frequencies detected in each observing season are given in 
Table \ref{freq1} and Table \ref{freq2}.

@@

Fig.~\ref{schematic} presents a schematic representation of GD358's 
independent frequencies detected between 1982 and 2016.  
The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret the bands to represent a series of modes of radial overtone $k$. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of stable multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands (Fig.~\ref{rot}). This does not imply that the 
structures found by \cite{Winget94}, \cite{Kepler03}, \cite{Montgomery10} and others are not real. In 
particular, it is clear that the 1741 \muHz\ ($k=12$) mode can be explained by the oblique rotation model during 2006. However, 
previous work represents short snapshots of GD358. When examined over timescales of decades, these structures 
do not regularly recur, and so cannot be interpreted as simple rotational splitting. The lower frequency band widths 
are most likely the result of various processes producing amplitude and/or frequency modulation that obscures any 
underlying signatures of rotational splitting. 
\begin{figure}
 \epsscale{1.0}
 \plotone{gd358k17_k8.eps}
% \plottwo{gd358multifk17.eps}{gd358multifk8.eps}
 \caption{The detailed schematic distribution of frequencies for the bands near 1300({\em{right}}
 and 2360 \muHz\ ({\em{left}}). As an example of the higher frequency bands, the 1300 \muHz\ band 
 shows no multiplet structure, unlike the 2360 \muHz\ band. Here we see clear evidence of multiplet 
 structure. Assuming the triplet represents
 $\ell=1$, this implies a rotation period of $\approx 1.5$ days. Note changes in the x scale for both axes.
 (A color version of this figure is available in the online journal.)
 \label{rot}}
\end{figure}
